---
title: "Assignment 3"
author: "Angela Huynh"
date: "2025-02-27"
output: html_document
---

# clear all plots: 
# try(dev.off(dev.list()["RStudioGD"]),silent=TRUE)
# try(dev.off(),silent=TRUE)

# clear environment:
# rm(list=ls())

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

library(ggplot2)
library(dplyr)
library(gridExtra)
library(scales)
data <- read.csv("diabetes.csv")

# Color-blind-friendly color palette
color_palette <- scale_fill_manual(values = c("#0072B2", "#D55E00"))

```{r A: Distribution (histogram, bar, density plot)}
# Distribution of Cholesterol levels

data <- read.csv("diabetes.csv")

plot1 <- ggplot(data, aes(x = chol)) +
  geom_histogram(binwidth = 19.4, fill = "#0072B2", color = "black") +
  theme_minimal() +
  labs(title = "Distribution of Cholesterol",
       x = "Frequency",
       y = "Cholesterol Level") +
  theme_minimal()

plot1 + geom_vline(aes(xintercept=mean(chol)),
            color="blue", linetype="dashed", size=1)

```
library(ggplot2)
install.packages(webr)
library(webr)
library(dplyr)
if(!require(devtools)) install.packages("devtools")
devtools::install_github("cardiomoon/moonBook")
devtools::install_github("cardiomoon/webr")

#----

install.packages("ggiraphExtra")  # Install ggiraphExtra
library(ggiraphExtra)
library(dplyr)

``` {r B: Proportion-based Donut Plot}

library(ggiraphExtra)
library(dplyr)

data <- read.csv("diabetes.csv")

data <- data %>% filter(frame != "") # removing empty values for frame

PD <- data %>%
  group_by(frame, gender) %>% #group_by to group by each unique combination of grame, gender; summarise 
  summarise(count = n(), .groups = 'drop') # summarise() applies the mean(), min(), or max(), etc.; in this case gets the n() no. of rows
  #.drop to make it a simple dataframe and not a grouped one

PieDonut(PD, aes(gender, frame, count = count), title = "Frame Distribution within Gender")




```


```{r C: Associations - include line of best fit and coefficient correlation for line; line can be nonlinear }
# Create a scatter plot
data <- read.csv("diabetes.csv")
ggplot(data, aes(x = age, y = chol)) +
  geom_point(color = "red", alpha = 0.7) +  # Points with color
  labs(title = "Cholestrol Levels against Age", 
       x = "Age", 
       y = "Cholestrol Levels") +
  theme_minimal() +
  geom_smooth(method = "lm") +
  geom_text(x=82, y=285, label="Correlation: 0.22")

cor(data$age, data$chol)

```


```{r D: Graph showing variable trend over age}

# Calculate the average Medical.Costs by age

data <- read.csv("diabetes.csv")

df_avg_glyhb <- data %>%
  group_by(age) %>%
  summarise(avg_gly = mean(glyhb, na.rm = TRUE))

# Create the line plot
ggplot(df_avg_glyhb, aes(x = age, y = avg_gly)) +
  geom_line(color = "darkgreen") +  # Line plot
  geom_point(color = "darkseagreen") +   # Optional: Points on the line to show data points
  labs(title = "Average Glycated Hemoglobin (Glyhb) by Age", 
       x = "Age", 
       y = "Glyhb") +
  theme_minimal()  # optional: cleaner plot theme

```

colors() # getting available colors in R


```{r E: Heamatp heatmap of chol, stab.glu, hdl, ratio, glyhb (y axis) as an increasing “age” of the patients in the x axis}

# Assuming data is already loaded from "diabetes.csv"
data <- read.csv("diabetes.csv")

# Rescale the variables before reshaping the data
data_scaled <- data %>% # scaled value represents the deviation from the mean in terms of standard deviations
  mutate(
    chol = scale(chol),
    stab.glu = scale(stab.glu),
    hdl = scale(hdl),
    ratio = scale(ratio),
    glyhb = scale(glyhb)
  )

# Reshape the data to long format for the heatmap
data_long <- data_scaled %>%
  select(age, chol, stab.glu, hdl, ratio, glyhb) %>%
  gather(key = "variable", value = "value", chol, stab.glu, hdl, ratio, glyhb)

# Create the heatmap
ggplot(data_long, aes(x = age, y = variable, fill = value)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", midpoint = median(data_long$value, na.rm = TRUE)) +
  theme_minimal() +
  labs(title = "Heatmap of Patient Data",
       x = "Age",
       y = "Variable",
       fill = "Scaled Value")
ggplot


```



```{r F: Ridgeline plot using “location” col for distributions of chol, stab.glu, hdl, ratio, and glyhb using separate graphs for each}

# install.packages("ggridges")
library(ggridges)

# Reshape the data to long format for ridgeline plots
data_ridgeline <- data %>%
  select(location, chol, stab.glu, hdl, ratio, glyhb) %>%
  gather(key = "variable", value = "value", chol, stab.glu, hdl, ratio, glyhb)


ggplot(data_ridgeline, aes(x = value, y = location, fill = location)) +
  geom_density_ridges() +
  facet_wrap(~ variable, scales = "free_x") +  # separate graphs for each variable
  theme_minimal() +
  labs(title = "Ridgeline Plot of Variables by Location",
       x = "Value",
       y = "Location")



```

